Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Conclusion.tex
stuff

>>>>

* Present work
- Abstract converted into ``have showns''
- Replacement policies/policies to decide when a request a sent etc (along with
  associativity/sizing of caches) is parameterized
* Future work
- Forward progress/liveness property
  - closely related to network/buffers
  - in presence of bounded buffers
  - in presence of communication fabric
- Automating Bluespec specification into Coq/vice versa
  - Verified compiler

### 2014-02-06 paper/Introduction.tex
stuff

The main contribution of this paper is \textbf{the first verification of a
cache coherence protocol that is parametric over an infinite family of
memory hierarchies}.  Our proof is implemented in the Coq proof assistant and
checked via a very small trusted code base.  We also want to draw attention to
a few key elements of our methodology that should be of interest for future
verifications of this kind, both through literal code reuse and through insights
about key invariants that should generalize across protocols.
\begin{enumerate}
\item We give a \emph{precise formal definition of a cache-coherent system} as one
that obeys \emph{store atomicity}. The definition is abstract in that it does not
rely on any hardware implementation.
\item We provide a \emph{Coq-based proof of the directory-based MSI cache coherence
protocol} with inclusive caches and no silent drops.
\item We present \emph{a set of invariants} that are not restricted to the MSI
protocol but which imply store atomicity. Thus, other protocols will also
satisfy store atomicity as long as they obey the invariants.
\item We present what we believe to be \emph{the key insight behind the
correctness of the MSI protocol}. We believe that this property is true also of
other protocols like MESI, MOSI, and MOESI, enabling reuse of portions of our
MSI proof for their verification.
\end{enumerate}
%In practice, verification proceeds in
%two steps. First, a cache coherence protocol is constructed and verified. This
%protocol generally directly represents many of the key concurrent aspects of an
%implementation (\eg{} reads and writes to the system are broken into local
%atomic micro-operations) but ignores implementation details (\eg{} buffer
%sizing and memory access times). Second, once the protocol has been verified,
%it is handed to hardware experts who fill in the missing implementation details
%and realize it at the Register Transfer Level (RTL) of gates, wires, and
%registers, via further breaking down of local micro-operations to make them
%realizable in hardware. This effort is not a one-time translation, as
%performance properties are not easily estimated; a significant amount of
%performance testing is requires to select a good implementation. 
%
%\adamc{I don't understand what this paragraph is saying.}
%From a methodological perspective, this two-step process leaves much to be
%desired. The significant effort of verification is effectively lost by the
%concurrent non-trivial tasks of translation from the protocol representation to
%a hardware description and the necessary refinement to add missing
%implementation details to the protocol.

>>>>

The main contribution of this paper is, \textbf{the first verification of a
cache coherence protocol that is parametric over an infinite family of
memory hierarchies}.  Our proof is implemented in the Coq proof assistant and
checked via a very small trusted code base.  We also want to draw attention to
a few key elements of our methodology that should be of interest for future
verifications of this kind, both through literal code reuse and through insights
about key invariants that should generalize across protocols.
\begin{enumerate}
\item We give a \emph{precise formal definition of a cache-coherent system} as one
that obeys \emph{store atomicity}. The definition is abstract in that it does not
rely on any hardware implementation.
\item We provide a \emph{Coq-based proof of the directory-based MSI cache coherence
protocol} with inclusive caches and no silent drops.
\item We present \emph{a set of invariants} that are not restricted to the MSI
protocol but which imply store atomicity. Thus, other protocols will also
satisfy store atomicity as long as they obey the invariants.
\item We present what we believe to be \emph{the key insight behind the
correctness of the MSI protocol}. We believe that this property is true also of
other protocols like MESI, MOSI, and MOESI, enabling reuse of portions of our
MSI proof for their verification.
\end{enumerate}
%In practice, verification proceeds in
%two steps. First, a cache coherence protocol is constructed and verified. This
%protocol generally directly represents many of the key concurrent aspects of an
%implementation (\eg{} reads and writes to the system are broken into local
%atomic micro-operations) but ignores implementation details (\eg{} buffer
%sizing and memory access times). Second, once the protocol has been verified,
%it is handed to hardware experts who fill in the missing implementation details
%and realize it at the Register Transfer Level (RTL) of gates, wires, and
%registers, via further breaking down of local micro-operations to make them
%realizable in hardware. This effort is not a one-time translation, as
%performance properties are not easily estimated; a significant amount of
%performance testing is requires to select a good implementation. 
%
%\adamc{I don't understand what this paragraph is saying.}
%From a methodological perspective, this two-step process leaves much to be
%desired. The significant effort of verification is effectively lost by the
%concurrent non-trivial tasks of translation from the protocol representation to
%a hardware description and the necessary refinement to add missing
%implementation details to the protocol.

### 2014-02-06 paper/RelatedWork.tex
stuff

For verification of high level specifications, modern industrial practice
consists of modeling small instances of the protocols, e.g., three CPUs
handling two addresses and one bit of data, in terms of interleaving atomic
steps, in guard/action languages such as Murphi \cite{tla} or TLA+ \cite{tla}, and
exploring the reachable states through explicit state enumeration.
Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
verification of cache coherence protocols.
Monolithic formal verification methods – methods that treat the protocol as a whole – have
been used fairly routinely for verifying cache coherence protocols from the early 1990s
[10 (2 nodes, murphi), 61 (), 141]. However, these monolithic techniques will not be able to handle the very
large state space of hierarchical protocols. Compositional verification techniques are essential.

>>>>

For verification of high level specifications, modern industrial practice
consists of modeling small instances of the protocols, e.g., three CPUs
handling two addresses and one bit of data, in terms of interleaving atomic
steps, in guard/action languages such as Murphi \cite{tla} or TLA+ \cite{tla}, and
exploring the reachable states through explicit state enumeration.
Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
verification of cache coherence protocols.
Monolithic formal verification methods – methods that treat the protocol as a whole – have
been used fairly routinely for verifying cache coherence protocols from the early 1990s
[10 (2 nodes, murphi)]. However, these monolithic techniques will not be able to handle the very
large state space of hierarchical protocols. Compositional verification techniques are essential.
